---
title: "Lab 4: Does Prenatal Care Improve Infant Health?"
author: "Carmen Easterwood, James Nguyen"
date: "April 23, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
load("bwght_w203.Rdata")
library(car)
library(caret)
library(corrplot)
library(gridExtra)
library(ggplot2)
library(lattice)
library(lmtest)
library(sandwich)
library(scales)
library(stargazer)
```

# Introduction

This report will analyze the relationship between prenatal care and health outcomes for newborn infants. Our team has been hired by a health advocacy organization (HAO) to study a sample of data from the Center for Disease Control's National Center for Health Statistics and from birth certificates, with a focus on helping the HAO understand the stated relationship. Our findings may be used to develop the HAO's advocacy agenda.

Since our primary goal is to understand the impact of prenatal care, we have divided our variables into the following categories:

+ Outcome (dependent) variables: `bwght`, `omaps`, `fmaps`
    + Outcomes that are a function of `bwght`: `lbw`, `vlbw`
+ Key explanatory (independent) variables: `monpre`, `npvis`
+ Additional covariates (independent variables):
    + Demographics: `mage`, `meduc`, `fage`, `feduc`, `male`, `mwhte`, `mblck`, `moth`, `fwhte`, `fblck`, `foth`
    + Vices: `cigs`, `drink`

As you read this report, keep the following caveats in mind:  

1. We cannot account for how our data sample was collected, so it is possible the data is not a random sample. We suspect it is not representative of the entire U.S. population.
2. There are some additional variables that may affect infant health, which we are unable to account for. These include, but are not limited to: the mother's diet, the quality of care received, and the presence of diseases that can be passed from mother to child (such as HIV or hemophilia).

The following sections document our analysis and findings.

# Exploratory Data Analysis

Our dataset has records for `r formatC(length(data$bwght), big.mark=",")` newborns, however some newborns are missing data for some variables. For our analysis, we will limit our dataset to only those newborns who have data available for every variable (i.e. complete cases). There are `r formatC(length(data[complete.cases(data),]$bwght), big.mark=",")` of these newborns.

```{r}
v <- na.exclude(data, complete.cases(v))
```

### Infant's Health

We have two methods of measuring the infant's health: by birthweight or by APGAR score. `bwght` is fairly normally distributed, with a mean weight of `r formatC(round(mean(v$bwght), 0), big.mark=",")` grams.

```{r}
ggplot(v, aes(x = bwght)) + geom_histogram(binwidth = 200) +
  ggtitle("Histogram of Infant Birthweights") + xlab("Birthweight (grams)") + ylab("# Infants")
```

Of the infants in our sample:

+ `r sum(v$lbw)` infants, or `r round(100*sum(v$lbw)/length(v$lbw), 1)`%, are considered "low birthweight", meaning they weigh less than 2,000 grams.
+ `r sum(v$vlbw)` infants, or `r round(100*sum(v$vlbw)/length(v$vlbw), 1)`%, are considered "very low birthweight", meaning they weigh less than 1,500 grams.

APGAR scores are an ordinal measure ranging from 0-10, and are used soon after birth to rate the health of newborns. We have data on our newborns' APGAR scores one minute and five minutes after birth. These variables are highly left-skewed, as most newborns have a score of 8 or 9 after one minute, and have moved to a score of 9 or even 10 after five minutes.

```{r}
plot.omaps <- ggplot(v, aes(x = omaps)) + geom_histogram(bins = 11) +
  ggtitle("Histograms of APGAR Scores") +
  scale_x_continuous(name = "1-Minute APGAR Score", breaks = seq(0, 10, by = 2)) +
  scale_y_continuous(name = "# Infants", limits = c(0, 1500))
plot.fmaps <- ggplot(v, aes(x = fmaps)) + geom_histogram(bins = 11) + ggtitle("") +
  scale_x_continuous(name = "5-Minute APGAR Score", breaks = seq(0, 10, by = 2)) +
  scale_y_continuous(name = "", limits = c(0, 1500))
grid.arrange(plot.omaps, plot.fmaps, ncol = 2)
```

### Prenatal Care

We have two ways of measuring prenatal care:

1. `monpre` is the month prenatal care began and ranges from 0 to 9.
2. `npvis` is the number of prenatal visits, and ranges from 0 to 40.

These variables are somewhat negatively correlated (`r round(cor(v$monpre, v$npvis), 3)`), since generally speaking, starting prenatal care earlier in pregnancy gives the mother more opportunities to make additional prenatal visits.

In our dataset, `monpre` is right skewed and most mothers have begun prenatal care by their second month of pregnancy. Most mothers went for 10-15 prenatal visits during their pregnancy, but the variable has a long right tail. Mothers with an extremely high number of prenatal visits likely had a high-risk pregnancy for one reason or another.

```{r}
plot.monpre <- ggplot(v, aes(x = monpre)) + geom_histogram(bins = 10) +
  scale_x_continuous(name = "Month Prenatal Care Began", breaks = seq(0, 9, by = 2)) +
  scale_y_continuous(name = "# Infants", limits = c(0, 750)) +
  ggtitle("Histograms for Prenatal Care") + ylab("# Infants")
plot.npvis <- ggplot(v, aes(x = npvis)) + geom_histogram(binwidth = 2) +
  scale_y_continuous(name = "", limits = c(0, 750)) +
  ggtitle("") + xlab("Number of Prenatal Visits")
grid.arrange(plot.monpre, plot.npvis, ncol = 2)
```

### Covariates

Our dataset includes additional covariates related to demographics and the mother's use of contraindicated substances ("vices") during pregnancy. Our vice variables describe the number of cigarettes and alcoholic drinks consumed by the mother during pregnancy. Both of these variables are highly right skewed, with large masses at 0. Note that for cigarettes, there seems to be some measurement error at the high end, as mothers appear to report the approximate number of packs rather than the exact number of cigarettes smoked (1 pack = 20 cigarettes). 

```{r}
plot.cigs <- ggplot(v, aes(x = cigs)) + geom_histogram(binwidth = 1) +
  scale_x_continuous(name = "Avg. Cigarettes per Day") +
  scale_y_continuous(name = "# Infants", limits = c(0, 1600)) +
  ggtitle("Histograms of Vice Variables")
plot.drink <- ggplot(v, aes(x = drink)) + geom_histogram(binwidth = 1) + ggtitle("") +
  scale_x_continuous(name = "Avg. Drinks per Week", breaks = seq(0, 9, by = 2)) +
  scale_y_continuous(name = "", limits = c(0, 1600))
grid.arrange(plot.cigs, plot.drink, ncol = 2)
```

We have the following information on the demographics of the mother, father, and baby:

+ `r sum(v$male)` infants, or `r round(100*sum(v$male)/length(v$male), 1)`%, are male.
+ The mother's age is quite normally distributed, is postively correlated with the father's age (+`r round(cor(v$mage, v$fage), 3)`), and has a median of `r median(v$mage)`. Meanwhile, the father's age has a slightly higher median at `r median(v$fage)`, and has a slightly longer right tail.

    ```{r}
    plot.mage <- ggplot(v, aes(x = mage)) +
      geom_histogram(binwidth = 2, fill = "red", alpha = 0.8) +
      scale_x_continuous(name = "Mother's Age", limits = c(15,65)) +
      scale_y_continuous(name = "# Infants", limits = c(0, 350)) +
      ggtitle("Histograms of Parent Age")
    plot.fage <- ggplot(v, aes(x = fage)) +
      geom_histogram(binwidth = 2, fill = "blue", alpha = 0.8) +
      scale_x_continuous(name = "Father's Age", limits = c(15,65)) +
      scale_y_continuous(name = "", limits = c(0, 350)) + ggtitle("")
    grid.arrange(plot.mage, plot.fage, ncol = 2)
    ```

+ The mother's and father's educations are also quite positively correlated (+`r round(cor(v$meduc, v$feduc), 3)`), and have similar distributions with spikes at 12 years (high school diploma), 14 years (associate's degree), and 16 years (bachelor's degree).

    ```{r}
    plot.meduc <- ggplot(v, aes(x = meduc)) +
      geom_histogram(binwidth = 1, fill = "red", alpha = 0.8) +
      scale_x_continuous(name = "Mother's Education (Years)", limits = c(3,17)) +
      scale_y_continuous(name = "# Infants", limits = c(0, 600)) +
      ggtitle("Histograms of Parent Education")
    plot.feduc <- ggplot(v, aes(x = feduc)) +
      geom_histogram(binwidth = 1, fill = "blue", alpha = 0.8) +
      scale_x_continuous(name = "Father's Education (Years)", limits = c(3,17)) +
      scale_y_continuous(name = "", limits = c(0, 600)) + ggtitle("")
    grid.arrange(plot.meduc, plot.feduc, ncol = 2)
    ```

+ The race of the mother and father is more complicated to analyze, since we have 3 indicator variables for each parent, which correspond to "white", "black", and "other" races. For ease of analysis, we combine the indicator variables into a single "race" variable for each parent, which allows us to view a table of the parents' races. Two things stand out in this table:

    1. Almost 90% of parents in this dataset are white.
    2. Over 95% of couples lie along the diagonal of the table, which means the mother and father are the same race.

    ```{r}
    mrace <- as.factor(c())
    levels(mrace) <- c("white", "black", "other")
    
    for (i in 1:length(v$bwght)){
      if (v[i,]$mwhte == 1) {mrace[i] <- "white"}
      else if (v[i,]$mblck == 1) {mrace[i] <- "black"}
      else if (v[i,]$moth == 1) {mrace[i] <- "other"}
    }
    
    frace <- as.factor(c())
    levels(frace) <- c("white", "black", "other")
    
    for (i in 1:length(v$bwght)){
      if (v[i,]$fwhte == 1) {frace[i] <- "white"}
      else if (v[i,]$fblck == 1) {frace[i] <- "black"}
      else if (v[i,]$foth == 1) {frace[i] <- "other"}
    }
    
    table(mrace, frace)
    ```

### Correlation Analysis

We have created a correlation matrix to identify pairs of variables that are highly correlated. Looking at the correlation matrix below, we can see that the following are highly correlated:

- Race of the baby's mother and father
- Ages of the baby's mother and father
- Education level of the baby's mother and father

```{r}
cor=cor(v)
corrplot(cor, method="circle", diag = FALSE, tl.col = "black")
```

# Model Specifications

### Outcome Variable

To evaluate the well-being of a newborn baby, we will use `bwght` as the only depedent variable. While APGAR scores are useful measures and could potentially be combined with weight to form a more comprehensive measure of well-being, we do not have the domain expertise to create such an index, and will just stick with `bwght`. With regard to well-being, we assume that heavier is better.

### Selection of Independent Variables

We will not include the variables `lbw` or `vlbw` in any model because these variables are a function of our outcome variable. In models 1 and 2 we will also exclude any variables that are highly correlated with each other, as determined in our **Correlation Analysis** section above. For example, since each of the pairs of mother and father race indicators (`mwhte`/`fwhte`, `mblck`/`fblck`, `moth`/`foth`) each have a correlation of ~90%, we will not include the race indicators for both parents in models 1 and 2.

### Model 1

Our first model contains only our variables of interest, which are the number of prenatal visits (`npvis`) and the month in which prenatal care began (`monpre`). We use robust standard errors in all of our models.

```{r}
model1 <- lm(bwght ~ monpre + npvis, data = v)
coeftest(model1, vcov = vcovHC)
summary(model1)$r.square
AIC(model1)
```
The coefficent variance test shows that `npvis` is highly statistically significant, although it is not practically significant. Each additional prenatal visit is associated with a birthweight increase of 13.4 grams for the newborn, which is about half an ounce.

Furthermore, our model's $R^2$ is very low. We need to include additional covariates to further explain the variation in newborn health, which we will do in our next model.

#### Assessment of CLM Assumptions for Model 1

+ CLM1 (Linear in the parameters): The population model can be written as $bwght = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + ... + \beta_k x_k + u$, where $\beta_0$, $\beta_1$, ..., $\beta_k$ are the unknown parameters (constants) of interest and $u$ is an unobserved random error or disturbance term. This is the base assumption as we fit our data to a linear model.

+ CLM2 (Random sampling): Our data comes from the CDC's National Center for Health Statistics, but it is not clear that it is a random sample. It is possible the data came from a certain community, since we see an unusually high percentage of white mothers (`r round(sum(v$mwhte)/length(v$mwhte)*100, 1)`%, whereas 61.6% of the general US population was white non-Hispanic in 2015). However, without further information, we will proceed as if this assumptions is met.

+ CLM3 (No perfect collinearity): This assumption is met. None of the variables are constant, and there is no perfect colinearity between `bwght`, `npvis`, and `monpre`. Furthermore, VIF scores for the independent variables are less than 4.

    ```{r}
    var(v$npvis)
    var(v$monpre)
    cor(v$npvis, v$monpre)
    vif(model1)
    ```

+ CLM4 (Zero conditional mean): The residuals vs. fitted plot below shows that the red smoothing curve (signifying the mean of the residuals) is very close to zero for most of the fitted values. Thus we can assume zero-conditional mean for our model.

    ```{r}
    plot(model1, which=1)
    ```

+ CLM5 (Homoskedasticity): Based on the residual vs. fitted plot, we cannot assume homoskedasticity. The width of the horizontal band increases at first, reaches a maximum around 3,400, then shrinks again. Also, we can run the formal Breusch-Pagan test to check against the null hypothesis of homoskedasticity, and we get a p-value smaller than 0.05, which means we can reject the null.

    ```{r}
    bptest(model1)
    ```

+ CML6 (Normality of errors): The distribution of residuals plot shows that the residual distribution is quite close to the normal distribution shape. Furthermore, the Q-Q plot shows that our error follows the normal distribution quite closely. Therefore we can assume that CLM6 is met for Model 1.

    ```{r}
    hist(model1$residuals, main ="Distribution of Model 1 Residuals", xlab = "Residual")
    plot(model1, which = 2)
    ```

### Model 2

In Model 2, we add variables that we believe will improve the accuracy of our prediction *without introducing bias*.

We add an indicator for the baby's gender and the variables that are related to the mother and her behavior: age, education, race, drinks per week, and cigarettes per day. We do not include information on the father's age, education, or race, since the mother's and father's information is highly correlated.

```{r}
model2 <- lm(bwght ~ monpre + npvis + male + cigs + drink + mage + magesq + meduc + mwhte +
               mblck, data = v)
coeftest(model2, vcov = vcovHC)
summary(model2)$r.square
AIC(model2)
```

In Model 2, `npvis` remains statistically significant and has a similar coefficient to Model 1. In addition, `male`, `cigs`, `mage`, `magesq`, `monpre`, and race are statistically significant. On average:

+ Male babies are 92 grams (3.2 ounces) heavier at birth than females.
+ An extra cigarette per day is associated with a 10-gram (0.35-ounce) reduction in birthweight. This is not practically significant, since a mother would need to smoke an additional 46 cigarettes *every day of her pregnancy* before we would expect to see a 1-pound reduction in weight.
+ The mother's age has a quadratic relationship with the weight of the baby. Birthweight is maximized when the mother is 31.7 years old, and declines at lower and higher ages. A 31.7-year-old mother's baby is 166 grams (5.9 ounces) heavier than a 20-year-old mother's baby, and 83 grams (2.9 ounces) heavier than a 40-year-old mother's baby.
+ White mothers give birth to babies that are 175 grams (6.2 ounces) heavier than mothers in the "other" race category. Black mothers give birth to babies that are 139 grams (4.9 ounces) heavier than "other" mothers, but this coefficient is only marginally statistically significant.

We are concerned that the coefficient on `monpre` is positive, which would indicate that babies benefit (in terms of higher birthweight) from starting prenatal care *later* in the pregnancy. This relationship clearly requires further study, and is potentially distorted by the correlation between `monpre` and `npvis`.

In any case, the $R^2$ on this model is only slightly higher than Model 1, and the AIC is only marginally lower, so this model is still missing some of the key factors in explaining a baby's birthweight.

#### Assessment of CLM Assumptions for Model 2

+ CLM1-CLM2: Same as model 1.

+ CLM3 (No perfect collinearity): None of the independent variables are constant, and we know from **Correlation Analysis** that there are no perfect correlations between any two variables. VIFs for all independent variables are less than 4, except for `mage` and `magesq`, which are obviously related to each other. Therefore, the CLM3 assumption is met.

    ```{r}
    var(v$npvis)
    var(v$monpre)
    var(v$male)
    var(v$cigs)
    var(v$drink)
    var(v$mage)
    var(v$meduc)
    var(v$mwhte)
    var(v$mblck)
    vif(model2)
    ```

+ CLM4 (Zero conditional mean): The residuals vs fitted plot shows that the red smoothing curve (signifying the mean of the residuals) is very close to zero for most of the fitted values. Thus, we can assume the CLM4 assumption is met.

    ```{r}
    plot(model2, which=1)
    ```

+ CLM5 (Homoskedasticity): The width of the horizontal band is fairly consistent across the range of the data in the scale-location plot below, although it is not perfect at the far end (possibly due to sparseness of the data). We can also run the formal Breusch-Pagan test to check against the null hypothesis of homoskedasticity, and we find the p-value is too large to reject the null hypothesis. We therefore assume that the homoskedasticity assumption is met.

    ```{r}
    plot(model2, which=3)
    bptest(model2)
    ```

+ CML6 (Normality of errors): The distribution of residuals plot shows that the residual distribution is quite close to the normal distribution shape. Furthermore, the Q-Q plot shows that our error follows the normal distribution quite closely. Therefore we can assume that CLM6 is met for Model 2.

    ```{r}
    hist(model2$residuals, main ="Distribution of Model 2 Residuals", xlab = "Residual")
    plot(model2, which =2)
    ```

### Model 3

In Model 3, we include additional covariates that are problematic for our model, primarily because they are highly correlated with some of the variables we added in Model 2.

```{r}
model3 <- lm(bwght ~ monpre + npvis + cigs + drink + mage + magesq + meduc + fage + feduc +
               male + mwhte + mblck + fwhte + fblck, data = v)
coeftest(model3, vcov = vcovHC) 
summary(model3)$r.square
AIC(model3)
```

The additional variables have caused an interesting change to our results. The majority of the coefficients stay roughly the same, and `fage` and `feduc` are not significant variables. However, the coefficients on `mwhte` and `mblck` have turned from positive to negative values, while the coefficients on the father's race indicators are highly significant and large positive numbers. The change in the mother's race coefficients can be explained by the fact that we included highly correlated variables in the model, which together distort our results. In almost all cases, the father's positive race coefficient will offset the mother's negative coefficient, since almost all mother-father pairs are the same race. See our section on CLM3 for additional discussion.

#### Assessment of CLM Assumptions for Model 3

- CLM1-CLM2: Same as models 1 and 2.
- CLM3 (No perfect collinearity): As shown in **Correlation Analysis**, there are no perfect correlations between any two variables, but the correlation between `fwhte`/`mwhte` and `fblck`/`mblck` are both nearly 0.9. The VIF scores below for these four variables are high due to their high correlations. Although they are not perfectly collinear, this model does have a multicollinearity problem.

    ```{r}
    cor(v$fwhte,v$mwhte)
    cor(v$fblck,v$mblck)
    vif(model3)
    ```

+ CLM4 (Zero conditional mean): The residuals vs fitted plot shows that the red smoothing curve (signifying the mean of the residuals) is very close to zero for most of the fitted values. Thus we can assume zero-conditional mean.
```{r}
plot(model3,which=1)
```

+ CLM5 (Homoskedasticity): The width of the horizontal band is fairly consistent in the middle, but tapers off at both ends. We can run the formal Breusch-Pagan test to check against the null hypothesis of homoskedasticity, and the p-value is small enough to reject the null, meaning this model shows some heteroskedasticity.
```{r}
plot(model3, which=3)
bptest(model3)
```

+ CML6 (Normality of errors): The distribution of residuals plot shows that the residuals are quite close to the normal distribution shape. Furthermore, the Q-Q plot shows that our error follows the normal distribution quite closely. Therefore we can assume that CLM6 is met for Model 3.

```{r}
hist(model3$residuals, main = "Distribution of Model 3 Residuals", xlab = "Residual")
plot(model3, which =2)
```

# Summary of Results

Please see the **Appendix** for a table showing our results for Models 1, 2, and 3. Model 1 is the simplest and only shows our two variables of interest, `monpre` and `npvis`, of which `npvis` is statistically but not practically significant. Model 2 adds additional covariates and uncovers some interesting relationships between `bwght` and other predictor variables, while model 3 adds even more covariates but results in distorted coefficients for some variables due to their high correlations with other predictor variables.

Unfortunately all of these models have a low $R^2$, which means we are lacking some key variables that could explain infant birthweights.

# Causality

In Model 1, Model 2, and Model 3, we pointed out that the assumptions CLM1, CLM2, CLM3 are met. In each model, `npvis` was also shown to be statistically significant, while `monpre` was either not significant or had a counterintuitive coefficient sign (possibly due to moderate correlation with `npvis`), so in this section we will focus on investigating `npvis`.

To further check the causality of `npvis` on baby weight, we test if `npvis` is correlated with the error term:

```{r}
cor(v$npvis, model1$residuals)
cor(v$npvis, model2$residuals)
cor(v$npvis, model3$residuals)
```

The numbers all come out to very near zero. This means manipulations to `npvis` do not impact the error term (i.e. $\frac{\partial u}{\partial(npvis)} = 0$), so we can assume that $\frac{\partial (bwght)}{\partial(npvis)} = \beta_{npvis}$. Therefore, we can say that `npvis` does have a positive causal effect on birthweight (note all `npvis` coefficients are positive).

One way we could go further to establish a causal relationship would be to run an experiment. However, in this context a true experiment would be unethical, because we would have to assign some women to receive no prenatal treatment.

### Direction of Biases

As noted previously, we believe there are other factors that may impact the weight of a newborn baby. These may include the mother's and father's income, their health condition (e.g. the mother's diet or the presence of specific conditions such as gestational diabetes), and the mother's quality of life during pregnancy (e.g. stress level), among others. Therefore, the coefficients on `npvis` and `monpre` may be biased in the above three models. We will demonstrate how this could be possible. For simplicity's sake, in this analysis we will consider only `npvis`, and ignore `monpre` and other covariates.

Now, let's assume the only missing independent variable from our model, as compared to a 'true' population model, is the mother's income (`in`). Thus the population model is:

$$ bwght = \beta_0 + \beta_{npvis} npvis + \beta_{in} in + u $$

Research has linked low incomes to low infant birthweights\footnote{https://www.guttmacher.org/journals/psrh/2007/infants-low-birth-weight-linked-low-income-mothers-chronic-stress}. This means we can safely assume that `in` has a positive effect on the baby's weight, i.e. $\beta_{in} > 0$.

Now we consider the regression of `npvis` on our missing variable `in`.

$$ npvis = \delta_0 + \delta_{in} in + \nu $$

Research has shown that lower-income pregnant women receive less prenatal care, experience higher levels of stress and are more likely to deliver premature babies than other women\footnote{http://www.apa.org/pi/ses/resources/publications/women.aspx}. Therefore we assume that the higher the mother's income, the higher the number of prenatal visits she makes during pregnancy, so $\delta_{in} > 0$.

So the bias on $\beta_{npvis}$ is $\beta_{in} \cdot \delta_{in}$, which is > 0. In other words, our current model, which is missing income, seems to overestimate the coefficient for `npvis`. Thus, while we find evidence that our model is causal, we must also consider that the coefficient on our prenatal care variable may be biased. This would mean we are incorrectly estimating the effect size.

We may also have a problem with certain variables absorbing the causal effect of prenatal care. In Model 1, we included the variable `monpre` in addition to `npvis`. This variable is negatively correlated with `npvis` (`r round(cor(v$npvis, v$monpre), 3)`). Since its coefficient is not significant and is close to zero, it does not introduce more bias into the model, but it does add variance in the beta estimation for `npvis`. In fact, if we just regress `bwght` on `npvis`, the standard error of the coefficient on `npvis` is reduced and the t-value is higher than in Model 1 (3.76 vs. 2.98). This shows that `monpre` is indeed absorbing some of the causal effect from `npvis`.

```{r}
model4 = lm(bwght ~ npvis, data = v)
coeftest(model4)
```

# Conclusion

Our investigation of the impact of prenatal care on infant health outcomes has revealed several interesting relationships (or lack thereof) and unique features of the data.

+ *Number of prenatal visits:* In all of our models, this variable was statistically, but not practically, significant. We suspect that our measures of prenatal care were not sufficient to measure the full impact of prenatal care, and further research should aim to include variables relating to the quality of care received and the specific tasks performed during the visits. Furthermore, mothers and babies with certain health conditions may need extra prenatal visits to manage those conditions, and this should be accounted for in the model.

+ *Highly correlated parents:* The mother's and father's demographics were highly correlated. Including both in the model caused distortions in our results, so we believe it is best to include only one parent's demographics.

+ *Omitted variables:* Our low $R^2$ values show that we are missing some key variables that would help explain the variation in `bwght`. Examples include the mother's income, diet, stress level, and quality of care received, among others. It is possible that these omitted variables have biased our results.

+ *Use of birth weight as the outcome variable:* Birthweight is an important part of infant health, but it is not the only possible measure, and may not even be the best measure. We were limited by our own knowledge, which does not extend to the use of ordinal variables (such as APGAR scores) as an outcome, as well as by our data, which doesn't include information on other possible measures such as length of gestation.

Overall, we do find a small amount of evidence that prenatal care can have a positive impact on infant birthweight, but recommend further research to get a fuller and more accurate picture of the effect size.

# Appendix

```{r}
se.model1 = sqrt(diag(vcovHC(model1)))
se.model2 = sqrt(diag(vcovHC(model2)))
se.model3 = sqrt(diag(vcovHC(model3)))
```

```{r, results='asis'}
stargazer(model1, model2, model3, type = "latex", header = FALSE,
          title = "Linear Models Predicting Infant Birthweight",
          se = list(se.model1, se.model2, se.model3),
          star.cutoffs = c(0.05, 0.01, 0.001))
```
